Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities

### 2023-10-26 paper/manuscript.Rmd
Refurbished Discussion

---
title: Food-web structure, stochasticity and response diversity drive temporal stability in complex communities
author: Alain Danet, Andrew P. Beckerman
output:
  bookdown::pdf_document2
always_allow_html: true
toc: false
bibliography: references.bib
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
---

@@

```{r, results = FALSE}
sapply(c("targets", "tidyverse", "magrittr", "cowplot", "here", "piecewiseSEM", "semEff", "ggcorrplot", "kableExtra"), require, character.only = TRUE)

@@

First, there are numerous definitions of community stability depending of the
stressors considered. For example, resistance and resilience are typically used
when a pulse perturbation is considered [@kefi_advancing_2019]. While in natural
settings where press and pulse perturbations are simultaneously present,
community variability measured as the coefficient of variation is typically used
[@thibaut_understanding_2013]. Defined at the community scale, community
variability can be partioned into population stability and asynchrony
[@thibaut_understanding_2013]. Synchronous population fluctuations will amplify
the negative effects of environmental fluctuations on temporal community
stability while compensatory dynamics and portefolio effects (i.e. asynchrony)
will stabilise temporal fluctuations of community biomass [@shoemaker_long_2022;
@zhao_biodiversity_2022]. Environmental forcing (such as food availability and
warming) and land-use intensity was reported to synchronise population
fluctuations [@benton_population_2001; @bluthgen_land_2016; @song_reduced_2015;
@liu_aridity_2023] while habitat degradation tends to decrease population
stability [@olivier_urbanization_2020]. Understanding the effects of global
change on community temporal stability therefore require to account for their
simultaneous effects on community and population scales [@kefi_advancing_2019;
@simmons_refocusing_2021].

Second, the relation between community stability and biodiversity in variable
environment is contingent on species richness per se, on the effects of species
richness on population stability and asynchrony, and ultimately
on the diversity of species responses to enviromental fluctuations [response
diversity, @ives_stability_2000; @elmqvist_response_2003]. Increasing
species richness has been reported to decrease population stability in
experimental grasslands [@tilman_biodiversity_2006; @roscher_identifying_2011],
while increasing population stability in natural settings
[@houlahan_negative_2018] [add mechanisms?]. In turn, higher species richness
often results in higher asynchrony, through portfolio effects (i.e. statistical
averaging effects due to independant species fluctuations) and compensatory
dynamics (due to species interactions and response
diversity of species due to the environment) [@loreau_species_2008;
@thibaut_understanding_2013]. Systematic increase in asynchrony with species
richness is the mechanism that drives reported positive relationships between
community stability and species richness [ref]. Anthropogenic pressures, such as
eutrophication, has been reported to weakens the relationship between species
richness and asynchony in grasslands [@hautier_eutrophication_2014] but such
empiricial evidence remains scarced. However, stressors such as biotic
homogenization [@buisson_toward_2013; @muthukrishnan_invasive_2020;
@saladin_rapid_2020] and environmental forcing that decrease species response
diversity should dampens the relationships between asynchrony and species
richness [@ives_stability_2000; @thibaut_understanding_2013] due to a decrease
in portefolio effects and compensatory effects. [Talk about the effects of
evenness in temporal variability and evenness in abundance? e.g.,
@zhao_biodiversity_2022; @thibaut_understanding_2013]. Because stressors can
affect both community stability per se and how community stability scales with
species richness, improving our understanding require to study both
[@ives_stability_2007; @hautier_anthropogenic_2015].

@@

Here we deliver inference from modelling of complex biodiverse communities on
how food-web structure (connectance, average interaction strength, average
trophic level), response diversity, and environmental stochasticity combine to
shape species richness-stability relationships.

@@

We added a stochastic natural mortality rate to the species dynamics,
representing a fluctuating environment. Based on the litterature and previous
models [@vasseur_environmental_2007], we assumed that
natural mortality rates scales inversely with the species body mass ($d_i =
d_0M_i^{-1/4}$, ), as all other physiological parameters. We used a basal
natural mortality rate of $d_0 = .4$, as done in previous
studies[@vasseur_environmental_2007].

We introduced environmental stochasticity by
considering stochastic variation of mortality rates. Doing so allows to consider
stochasticity without modifying species interactions themselves
[@vasseur_environmental_2007]. The stochastic part of the mortality rates
($\epsilon_{i, t}$) follow a normal distribution 0 centered
($\epsilon_{i, t} \sim N(0,\sigma^2_e)$) with a variance ($\sigma^2_e$). The mortality rate each
species at the time $t$ equals to $d_{i,t} = d_i e^{\epsilon_{i, t}}$. We did so
to ensure to not have negative mortality rates [@vasseur_environmental_2007]. We
simulated $\epsilon$ using an Ornstein-Uhlenbeck process, which is a modified
browian motion where the stochastic values tend to come back to the central
value (i.e. $\epsilon = 0$) such as $\epsilon_t$ varies according to this
stochastic differential equation: $d\epsilon_t = (0 - \epsilon_t)d_t + \sigma_e
dW_t$, where $dW_t$ is a Browian motion. The strength of environmental
stochasticity was then controlled by $\sigma_e$.

@@

To vary interaction strength and mortality rates in the community,
we varied the Predator-Prey Bodysize Ratio (PPBR) from 1 to 100. We further
varied the functional reponse of consumer from type II to type III (resp. $h=2$
and $h=3$, eq XX) as it was shown to have a tremendous role on stability
[@brose_allometric_2006]. We set $K = XX$ to ensure that consumers were not
limited by the biomass input from primary producers.

@@

To further get a mechanistic insight about the effects of stressors and food-web
structure on stability, we partionned asynchrony into statistical averaging
effect (SAE) and compensatory dynamics effect (CPE) [@zhao_biodiversity_2022].
SAE quantities the portfolio effect emerging from independant species
fluctuations, while CPE describes compensatory dynamics emerging from species
interactions and differential responses to environmental stochasticity.

@@

Statistical averaging effect is the share of asynchrony assuming that species are
independant, meaning that the variance total of the community is equal to the
sum of species variances only (i.e. species covariances are null). We can
define community stability ($S_{com, IP}$) in that scenario and then derive $SAE$.

@@

As SAE is further dependant on the evenness of species variance, we
partionned $SAE$ into $SAE_{even}$ and evenness $EVN$, such as $SAE_{even} =
\sqrt{S}$ and $EVN = SAE / SAE_{even}$.

Compensatory dynamics in the food-webs is then the remaining part of asynchony:

@@

Compensatory dynamics can arise from predator-prey interactions
($CPE_{int}$) or from differential response of species to enviromental
stochasticity ($CPE_{env}$). From the deterministic and stochastic parts of
mortality rates, we computed the compensatory dynamics linked to the environment
such as:

@@

Using a structural equation model, we disantengled the complex effects of
food-web structure, environmental stochasticity and response diversity on the
different facets of temporal stability of biomass (i.e. population stability,
statistical averaging effects and compensatory dynamic effects). We set the
paths of the path of the model according to theorical expectations. We tested
the effects of initial species richness, initial connectance, Predator-Prey size
ratio on secondary structure of the food-webs, i.e. realized species richness
and connectance, the average trophic level, the average omnivory and the
interaction strength. Further, we expected those food-web structure to alter
population stability and synchrony. Grounded in theoretical ecology, we expected
that higher average trophic level, average omnivory, lower interaction strength,
and type III functional response (relative to type II) increase population
stability [@mccann_weak_1998; @mccann_reevaluating_1997;
@shanafelt_stability_2018]. Higher Predator-Prey size ratio was also expected to
increase population stability [@brose_allometric_2006]. Finally, the effect of
enviromental stochasticity was assumed to be mainly associated to population
stability (i.e. by increasing amplitude of population fluctuations).

@@

In a second analysis, we aimed to test how stability-species richness
relationships were modulated by food-web structure and the two stressors:
environmental stochasticity and response diversity. We modelled temporal
stability of biomass according to food-web structure: species richness, weighted
average trophic level, connectance, hill exponent, predator-prey size ratio, and
average interaction strength, as well as enviromental stochasticity and species
cross-correlation. We further included the two way interactions between species
richness and food-web structure, plus between species richness and stressors.
Finally we added the three way interaction between species richness, food-web
structure and stressors. The VIF were inferior to 5, indicating low
multicollinearity.

@@

```{r}
#TODO: change most of it for simplified SEM
tar_load(c(sem_eff_tab, sem_simplified_eff_tab))

@@

We further found that food-web structure had a strong impact on statistical
averaging effects, which raises from independant species fluctuations weigthed
the evenness of their variance. As expected the mathematical relationship (eq.
XX), species richness was a main positive driver of statistical averaging effect
($r_\delta$= `r rsr`, Fig. \@ref(fig:sem)), thereby increasing asynchrony.
Average trophic level and average interaction strength had strong negative
effects on statistical averaging effects (resp. $r_\delta$= `r rst` and
`r rsi`), thereby downplaying the portefolio effect raised from species
richness. In turn, we found that average interaction strength had a small
positive effect on compensatory dynamics ($r_\delta$= `r rci`, Fig.
\@ref(fig:sem)). Cross-correlation among species mortality rates (i.e. lower
response diversity) had a strong negative effect on compensatory dynamics
($r_\delta$= `r rco`), meaning that the loss of response diversity result in a
strong decrease in compensatory dynamics.

@@

```{r}
semtot_cap <- "Total effect of food-web structure on community stability,
asynchony, and population stability. The total effects were derived from the
structural equation model."
```

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both enviromental stochasticity and cross-correlation among
species had by far the most negative effect on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsrho`, Fig. \@ref(fig:semtot)). Species richness
had a low overall positive total effect on community stability (`r rtsr`)
because its strong total positive effect on asynchony (`r rtar`) was downplayed
by its total negative effect on population stability (`r rtpr`) and that the
weigtht of population stability on community stability is twice higher than the
one of asynchony. Average trophic level, type III functional response and
predator prey bodysize ratio had positive effects on community stability in
total (resp. `r rtst`, `r rtsh`, `r rtsz`, Fig. \@ref(fig:semtot)), driven by
their positive effects on population stability that were weakly counterbalanced
by their negative effects on asynchrony. In turn, connectance and average
interaction strength had low total negative effects on community stability (resp.
`r rtsc` and `r rtsi`, Fig. \@ref(fig:semtot)). While connectance had negative
effects both on population stability and asynchrony, average interaction
strength had a low total positive effect on population stability counterbalanced
by a stronger negative effect on asynchrony.

@@

Using a more complex linear model (i.e. including interactions), we investigated more
in detail the relationship between community stability and species richness and
how this relation is affected by environmental stochasticity and cross
correlation among species mortality rates. We found that environmental
stochasticity decrease overall community stability but has only little effect on
the shape of the community stability-species richness relationship (put some
numbers, Fig. \@ref(fig:pred)a and b). In contrasts, cross correlation among
species mortality rates both decreases overall community stability and nullifies
the community stability-species richness relationship. Surpringly, we found that
despite average interaction strength has a small overall effect on
community stability, high average interaction strength change the shape of the
community stability-species richness relationship from positive to negative when
response diversity is low (put some numbers, Fig. \@ref(fig:pred)b).

@@

## Food-web structure and stressors affects population stability and asynchony

Our results shows that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, confirming previous results in
empirical data [@danet_species_2021]. We further found that average trophic
level and functional response of type III increased average population
stability, confirming previous results [@brose_allometric_2006;
@shanafelt_stability_2018; @danet_species_2021;
@eschenbrenner_diversity_nodate]. Higher average trophic level stabilises
populations because higher predator are themselves more stable and because they
control the amplitude of the fluctuations of their prey [REF, we can test
that?]. Furthermore, average omnivory was strongly correlated with average
trophic level, suggesting that average omnivory also stabilise population
stability, in line with previous theoretical results in small modules
[@mccann_reevaluating_1997]. In turn, functional response of type III decreases
the feeding rates of consumers when their preys have a low biomass which leads
to a less dramatic fluctuations of prey biomass [@uszko_effects_2017;
@brose_allometric_2006]. We however found no big effect of species richness on
population stability, which has been reported to be negative, positive or
neutral in previous studies [@ruijven_contrasting_2007;
@houlahan_negative_2018].

Our results shows that asynchrony is mainly decreased by the cross-correlation
among species mortality rates and mediated by compensatory dynamics, indicating
that the loss of response diversity suppressed the positive effects of species
richness on stability. Our result shows that evenness of biomass plays an
significant role in driving statistical averaging effect (i.e. portefolio
effects). Our results shows that average trophic level decreases strongly the
evenness of biomass variability, thereby limiting the effect of species richness on the
portefolio effects. Furthermore, increase in species richness also resulted in
less even biomass variability. Our results also suggest that stronger species
interaction, higher generalism, and functional response of type III are
associated with a more even variability, thereby raising the portefolio
effects. The importance of evenness in driving portefolio effects has been
recognised early by theoretical studies [@thibaut_understanding_2013], but
rather in terms of abundance than in variability [@zhao_biodiversity_2022].
We suggest that higher average trophic level decreases evenness of variability
because the model assumes that species of higher trophic level have a lower
overall mortality rates, similarly to previous studies [@vasseur_environmental_2007].

Our results further suggest that higher average trophic level results in lower
compensatory dynamics raising from species interactions. It might be explained
by the fact that average trophic level is strongly associated with average
omnivory, which might tamper synchronized predator-prey oscillations that can
raise compensatory dynamics [@mccann_reevaluating_1997]. Higher generalism in
also drives lower compensatory dynamics due to species interaction and then lend
to the same hypothesis such as more trophic links per consumer should decrease
overall amplitude of oscillations due to prey switching. Furthermore, functional
response of type III which drives prey switching at low abundance of the
prefered prey also lend toward this interpretation
[@mccann_diversitystability_2000]. [Do not have clear idea for richness at this
time]. Surpringly however, average interaction strength had no effects on
compensatory dynamics that is not already captured by other food-web structural
properties. Compensatory dynamics due to environment was decreased by both
species richness and cross-correlation among species mortality rates (i.e. loss
of response diversity). Our interpretation is that higher species richness
syncronize more species fluctuations due to mortality because each added species
adds (S - 1, S being richness) positive covariance terms in the stochastic
mortality rates, thereby increasing the overall synchrony in mortality rates
when cross correlation are not null.

## Stochasticity and response diversity affect stability-richness relationship

Our results suggest that, confirming the result of the structural equation
model, that in maximal response diversity environmental stochasticity primarily
drives average stability, without changing the shape of the stability-richess
relationship. The reason is that the strength of environmental stochasticity
only affect population stability and not asynchrony, the latter being the driver
of species-richness relationships. In turn, a decrease in response diversity
dampens the stability-richness relationship [@thibaut_understanding_2013].
Interestingly, our results also suggest that high response diversity also
decrease the destabilising effects of average interaction strength on
stability-richness relationships. Indeed, in low response diversity, we found
high interaction strength leads to negative stability-richness relationships.

## Implication for real world

Our results highlight that population stability is the primary driver of
community stability. In the context of global changes in which environmental
variability is expected to increase tremendously, we expect to see a strong
decrease in community stability, despite high response diversity. But ongoing
biotic homogenization [@olden_homogocene_2018; @muthukrishnan_invasive_2020;
@wang_biotic_2021] is also suggested to descrease response diversity
[@wang_biotic_2021; @su_human_2021], which in turn is expected to further
decrease stability despite high species richness, even more in food-web
communities composed of high interaction strength. Then, our study reinforces
that the restauration of ecosystems along with a restauration of landscape
complexity that provides refuges from extrem environment are key to maintain
temporal stability of ecological communities [@ryser_landscape_2021].

# Conclusion

TODO

>>>>

---
title: Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities
author: Alain Danet, (Sonia Kéfi, Thomas F Johnson), Andrew P. Beckerman
output:
  officedown::rdocx_document:
    base_format: "bookdown::word_document2"
    number_sections: false
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
always_allow_html: true
bibliography: references.bib
link-citations: true
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

```{r, results = FALSE}
sapply(c("targets", "tidyverse", "magrittr", "cowplot", "here", "piecewiseSEM",
    "semEff", "ggcorrplot", "kableExtra", "officedown"),
  require,
  character.only = TRUE)

@@

First, there are numerous definitions of community stability depending of the
stressors considered. For example, resistance and resilience are typically used
when a pulse perturbation is considered [@kefi_advancing_2019]. While in natural
settings where press and pulse perturbations are simultaneously present,
community variability measured as the coefficient of variation is typically used
[@thibaut_understanding_2013]. Defined at the community scale, community
variability can be partitioned into population stability and asynchrony
[@thibaut_understanding_2013]. Synchronous population fluctuations will amplify
the negative effects of environmental fluctuations on temporal community
stability while compensatory dynamics and portfolio effects (i.e. asynchrony)
will stabilise temporal fluctuations of community biomass [@shoemaker_long_2022;
@zhao_biodiversity_2022]. Environmental forcing (such as food availability and
warming) and land-use intensity was reported to synchronise population
fluctuations [@benton_population_2001; @bluthgen_land_2016; @song_reduced_2015;
@liu_aridity_2023] while habitat degradation tends to decrease population
stability [@olivier_urbanization_2020]. Understanding the effects of global
change on community temporal stability therefore require to account for their
simultaneous effects on community and population scales [@kefi_advancing_2019;
@simmons_refocusing_2021].

Second, the relation between community stability and biodiversity in variable
environment is contingent on species richness per se, on the effects of species
richness on population stability and asynchrony, and ultimately
on the diversity of species responses to environmental fluctuations [response
diversity, @ives_stability_2000; @elmqvist_response_2003]. Increasing
species richness has been reported to decrease population stability in
experimental grasslands [@tilman_biodiversity_2006; @roscher_identifying_2011],
while increasing population stability in natural settings
[@houlahan_negative_2018] [add mechanisms?]. In turn, higher species richness
often results in higher asynchrony, through portfolio effects (i.e. statistical
averaging effects due to independent species fluctuations) and compensatory
dynamics (due to species interactions and response
diversity of species due to the environment) [@loreau_species_2008;
@thibaut_understanding_2013]. Systematic increase in asynchrony with species
richness is the mechanism that drives reported positive relationships between
community stability and species richness [ref]. Anthropogenic pressures, such as
eutrophication, has been reported to weakens the relationship between species
richness and asynchrony in grasslands [@hautier_eutrophication_2014] but such
empirical evidence remains scarce. However, stressors such as biotic
homogenization [@buisson_toward_2013; @muthukrishnan_invasive_2020;
@saladin_rapid_2020] and environmental forcing that decrease species response
diversity should dampens the relationships between asynchrony and species
richness [@ives_stability_2000; @thibaut_understanding_2013] due to a decrease
in portfolio effects and compensatory effects. [Talk about the effects of
evenness in temporal variability and evenness in abundance? e.g.,
@zhao_biodiversity_2022; @thibaut_understanding_2013]. Because stressors can
affect both community stability per se and how community stability scales with
species richness, improving our understanding require to study both
[@ives_stability_2007; @hautier_anthropogenic_2015].

@@

Here we deliver inference from modelling of complex biodiverse communities on
how food-web structure (connectance, average interaction strength, average
trophic level), response diversity, and environmental stochasticity combine to
shape temporal stability of community biomass, as well as its partition into
asynchrony and population stability. We then assess how food-web complexity and
response diversity shape stability-richness relationships.

@@

We added a stochastic natural mortality rate to the species dynamics,
representing a fluctuating environment. Based on the literature and previous
models [@vasseur_environmental_2007], we assumed that
natural mortality rates scales inversely with the species body mass ($d_i =
d_0M_i^{-1/4}$, ), as all other physiological parameters. We used a basal
natural mortality rate of $d_0 = .4$, as done in previous
studies[@vasseur_environmental_2007].

We introduced environmental stochasticity by
considering stochastic variation of mortality rates. Doing so allows to consider
stochasticity without modifying species interactions themselves
[@vasseur_environmental_2007]. The stochastic part of the mortality rates
($\epsilon_{i, t}$) follow a normal distribution 0 centered
($\epsilon_{i, t} \sim N(0,\sigma^2_e)$) with a variance ($\sigma^2_e$). The mortality rate each
species at the time $t$ equals to $d_{i,t} = d_i e^{\epsilon_{i, t}}$. We did so
to ensure to not have negative mortality rates [@vasseur_environmental_2007]. We
simulated $\epsilon$ using an Ornstein-Uhlenbeck process, which is a modified
browian motion where the stochastic values tend to come back to the central
value (i.e. $\epsilon = 0$) such as $\epsilon_t$ varies according to this
stochastic differential equation: $d\epsilon_t = (0 - \epsilon_t)d_t + \sigma_e
dW_t$, where $dW_t$ is a Brownian motion. The strength of environmental
stochasticity was then controlled by $\sigma_e$.

@@

To vary interaction strength and mortality rates in the community,
we varied the Predator-Prey Bodysize Ratio (PPBR) from 1 to 100. We further
varied the functional response of consumer from type II to type III (resp. $h=2$
and $h=3$, eq XX) as it was shown to have a tremendous role on stability
[@brose_allometric_2006]. We set $K = XX$ to ensure that consumers were not
limited by the biomass input from primary producers.

@@

To further get a mechanistic insight about the effects of stressors and food-web
structure on stability, we partitioned asynchrony into statistical averaging
effect (SAE) and compensatory dynamics effect (CPE) [@zhao_biodiversity_2022].
SAE quantities the portfolio effect emerging from independent species
fluctuations, while CPE describes compensatory dynamics emerging from species
interactions and differential responses to environmental stochasticity.

@@

Statistical averaging effect is the share of asynchrony assuming that species are
independent, meaning that the variance total of the community is equal to the
sum of species variances only (i.e. species covariances are null). We can
define community stability ($S_{com, IP}$) in that scenario and then derive $SAE$.

@@

As SAE is further dependant on the evenness of species variance, we
partitioned $SAE$ into $SAE_{even}$ and evenness $EVN$, such as $SAE_{even} =
\sqrt{S}$ and $EVN = SAE / SAE_{even}$.

Compensatory dynamics in the food-webs is then the remaining part of asynchrony:

@@

Compensatory dynamics can arise from predator-prey interactions
($CPE_{int}$) or from differential response of species to environmental
stochasticity ($CPE_{env}$). From the deterministic and stochastic parts of
mortality rates, we computed the compensatory dynamics linked to the environment
such as:

@@

Using a structural equation model, we disentangled the complex effects of
food-web structure, environmental stochasticity and response diversity on the
different facets of temporal stability of biomass (i.e. population stability,
statistical averaging effects and compensatory dynamic effects). We set the
paths of the path of the model according to theoretical expectations. We tested
the effects of initial species richness, initial connectance, Predator-Prey size
ratio on secondary structure of the food-webs, i.e. realized species richness
and connectance, the average trophic level, the average omnivory and the
interaction strength. Further, we expected those food-web structure to alter
population stability and synchrony. Grounded in theoretical ecology, we expected
that higher average trophic level, average omnivory, lower interaction strength,
and type III functional response (relative to type II) increase population
stability [@mccann_weak_1998; @mccann_reevaluating_1997;
@shanafelt_stability_2018]. Higher Predator-Prey size ratio was also expected to
increase population stability [@brose_allometric_2006]. Finally, the effect of
environmental stochasticity was assumed to be mainly associated to population
stability (i.e. by increasing amplitude of population fluctuations).

@@

In a second analysis, we aimed to test how stability-species richness
relationships were modulated by food-web structure and the two stressors:
environmental stochasticity and response diversity. We modelled temporal
stability of biomass according to food-web structure: species richness, weighted
average trophic level, connectance, hill exponent, predator-prey size ratio, and
average interaction strength, as well as environmental stochasticity and species
cross-correlation. We further included the two way interactions between species
richness and food-web structure, plus between species richness and stressors.
Finally we added the three way interaction between species richness, food-web
structure and stressors. The VIF were inferior to 5, indicating low
multicollinearity. We then used the coefficients of the model to predict the shape
of the relationship between community stability and species richness according
to a gradient of response diversity and of food-web structure.

@@

```{r}
tar_load(c(sem_eff_tab, sem_simplified_eff_tab))

@@

We further found that food-web structure had a strong impact on statistical
averaging effects, which raises from independent species fluctuations weighted
the evenness of their variance. As expected the mathematical relationship (eq.
XX), species richness was a main positive driver of statistical averaging effect
($r_\delta$= `r rsr`, Fig. \@ref(fig:sem)), thereby increasing asynchrony.
Average trophic level and average interaction strength had strong negative
effects on statistical averaging effects (resp. $r_\delta$= `r rst` and
`r rsi`), thereby downplaying the portfolio effect raised from species
richness. In turn, we found that average interaction strength had a small
positive effect on compensatory dynamics ($r_\delta$= `r rci`, Fig.
\@ref(fig:sem)). Cross-correlation among species mortality rates (i.e. lower
response diversity) had a strong negative effect on compensatory dynamics
($r_\delta$= `r rco`), meaning that the loss of response diversity result in a
strong decrease in compensatory dynamics.

@@

```{r}
semtot_cap <- "Total effect of food-web structure on community stability,
asynchrony, and population stability. The total effects were derived from the
structural equation model."
```

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both environmental stochasticity and cross-correlation among
species had by far the most negative effect on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsrho`, Fig. \@ref(fig:semtot)). Species richness
had a low overall positive total effect on community stability (`r rtsr`)
because its strong total positive effect on asynchrony (`r rtar`) was downplayed
by its total negative effect on population stability (`r rtpr`) and that the
weight of population stability on community stability is twice higher than the
one of asynchrony. Average trophic level, type III functional response and
predator prey bodysize ratio had positive effects on community stability in
total (resp. `r rtst`, `r rtsh`, `r rtsz`, Fig. \@ref(fig:semtot)), driven by
their positive effects on population stability that were weakly counterbalanced
by their negative effects on asynchrony. In turn, connectance and average
interaction strength had low total negative effects on community stability (resp.
`r rtsc` and `r rtsi`, Fig. \@ref(fig:semtot)). While connectance had negative
effects both on population stability and asynchrony, average interaction
strength had a low total positive effect on population stability counterbalanced
by a stronger negative effect on asynchrony.

@@

Using a more complex linear model (i.e. including interactions), we investigated more
in detail the relationship between community stability and species richness and
how this relation is affected by environmental stochasticity and cross
correlation among species mortality rates. We found that environmental
stochasticity decrease overall community stability but has only little effect on
the shape of the community stability-species richness relationship (put some
numbers, Fig. \@ref(fig:pred)a and b). In contrasts, cross correlation among
species mortality rates both decreases overall community stability and nullifies
the community stability-species richness relationship. Surprisingly, we found that
despite average interaction strength has a small overall effect on
community stability, high average interaction strength change the shape of the
community stability-species richness relationship from positive to negative when
response diversity is low (put some numbers, Fig. \@ref(fig:pred)b).

@@

## Food-web structure and stressors affects population stability and asynchrony

Our results show that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, confirming previous results in
empirical fish communities [@danet_species_2021], which contrasts with empirical
data in birds and bats assemblages that reported a equivalent influence of
population stability and asynchrony on community stability
[@olivier_urbanization_2020]. Although the number of empirical and theoretical
studies is still limited to compare the respective contribution of asynchrony and
population stability to community stability in competitive assemblages and
food-webs, we suggest two non-exclusive and complementary explanations that
might drive a lower contribution of asynchrony in food-webs compared to
competitive assemblages. First, the range of bodymasses is much higher in
food-webs, which typically include species that differ by several order of
magnitude [@brose_predator_2019], which in turn creates highly uneven biomass
distribution, thereby creating uneven contributions of species to asynchrony
[@thibaut_understanding_2013] and limits the raise of high portfolio effects and
compensatory dynamics in species rich communities [@zhao_biodiversity_2022].
Second, the trophic interactions characterising food-webs create strong
interdependence among species, which in turn limits the raise of independent
fluctuations (i.e. portfolio effects) and compensatory dynamics. The question of
which mechanisms and community structure features drive differences in the
contribution of asynchrony and population stability in food-webs and competitive
assemblages is still a open question that might be crucial to accurately
address diagnostics and conservation measures. On top of food-web structure,
we found that environmental stochasticity and environmental correlation among
species was the main drivers of community stability, suggesting that stressors
are the first source of stability. This result is in accordance with previous
studies showing strong effects on environmental stochasticity
[@vasseur_environmental_2007] and the loss of response diversity decrease
community stability in tri-trophic food-webs [@ives_stability_2000;
@thebault_trophic_2005].

We further found that average trophic level increased average population
stability, confirming previous theoreical [@brose_allometric_2006;
@shanafelt_stability_2018; @eschenbrenner_diversity_nodate] and
[@danet_species_2021] empirical results. Higher average trophic level can
stabilise populations because higher predators  are themselves more stable due
to lower mortality rates with in turn makes them less sensible stochastic
perturbation and because they control the amplitude of the fluctuations of their
prey [@shanafelt_stability_2018, we can test that?]. In turn, we found that that
average trophic level decreases strongly asynchrony through a negative effect on
statistical averaging effect. Further analysis showed that average trophic level
limited statistical averaging effect by decreasing the evenness of biomass
variability. It is linked to the model assumption that species of higher trophic
level have a lower overall mortality rates, similarly to previous studies
[@vasseur_environmental_2007], stabilising populations and simultaneously
synchronising them. The fact that Predator Size Mass Ratio displays the same
effect, thereby building food-webs with bigger species having lower mortality
rates, further lends toward this hypothesis. Furthermore, average omnivory was
strongly correlated with average trophic level, suggesting that average omnivory
also stabilise population stability, in line with previous theoretical results
in small modules [@mccann_reevaluating_1997]. Our results further suggest that
higher average trophic level results in lower compensatory dynamics raising from
species interactions. It might be explained by the fact that average trophic
level is strongly associated with average omnivory, which might tamper
synchronized predator-prey oscillations that can raise compensatory dynamics
[@mccann_reevaluating_1997].

We however found no big effect of species richness on population stability,
which has been reported to be negative, positive or neutral in previous studies
[@ruijven_contrasting_2007; @houlahan_negative_2018]. [interpretation?].
However, species richness had a strong positive on asynchrony, by raising
statistical averaging effects. [add blabla about richness asynchrony]. In turn,
we found that species richness magnified the effect of the loss of response
diversity on the compensatory dynamics due to environment. Our interpretation is
that higher species richness synchronizes more species fluctuations due to
environmental mortality when species are not independent because each added
species adds (S - 1, S being richness) positive covariance terms in the
stochastic mortality rates, thereby increasing the overall synchrony in
mortality rates when cross correlation are not null. This assumption that each
species are equally correlated to each other seems unlikely to be true but as
far we know, there is compelling evidence yet for alternative correlation
structures. Connectance, quantifying generalism in food-web, had negative
effects on both asynchrony and population stability. We further found that
higher connectance leaded to less even biomass variability, thereby limiting the
raise of the portfolio effects, and leaded to lower compensatory dynamics due
to species interactions. This results in accordance with empirical finding that
a generalist predator can indeed synchronises the dynamics of a large number of
preys [@raimondo_interspecific_2004]. [No idea about connectance decreasing
population stability]. However, the overall effect of connectance on overall
stability was quite low, similarly to previous empirical findings
[@danet_species_2021], but in apparent disagreement with the predominant role of
connectance on stability in theoretical studies [complexity-stability paradox,
@may_will_1972].

Food-web dynamical features, functional response of type III and average
interaction strength, also had an effect on community stability. Our results
confirms the stabilising effect of the functional response of type III on
populations [@uszko_effects_2017; @brose_allometric_2006]. Compared to type II,
functional response of type III models a decrease of the feeding rates of
consumers when their preys have a low biomass which leads to a less dramatic
fluctuations of prey biomass. Furthermore, we found that functional response of
type III decreased asynchrony by decreasing compensatory dynamics due to species
interaction. This effect might be due to prey switching of consumers when one of
them reach low abundance [@mccann_diversitystability_2000]. Average interaction
strength, as connectance, had overall little effect on community stability.
Average interaction strength was found to decrease asynchrony by decreasing the
evenness of biomass variability, thereby decreasing the portfolio effect,
despite having a small positive effect on compensatory dynamics. In turn, we
surprisingly found that average interaction strength had little effects on
population stability, which contrasts with the importance of this parameter in
complexity-stability relationship [@may_will_1972].

We found that environmental stochasticity does not affect the shape of the
community stability - richness but rather control the overall community
stability (i.e. the intercept of the relation). At the opposite, In turn, a
decrease in response diversity dampens the stability-richness relationship. The
reason is that the strength of environmental stochasticity only affect
population stability and not asynchrony, the latter being the driver of
species-richness relationships [@ives_stability_2000; @thebault_trophic_2005;
@thibaut_understanding_2013]. Average interaction strength and connectance that
were found to have little effects on overall stability, were found to have a
strong effect on stability-richness relationships. This result is in accordance
with the complexity-stability theory that predicts that community stability
(i.e. measured as the dominant eigen value of the jacobian matrix) should
decrease as the community is species diversity, has higher connectance, and
higher interaction strength [@may_will_1972]. However, we found that the
negative effects of complexity on stability-richness relationship are cancelled
when response diversity is high. A interpretation is that environmental
stochasticity itself does not override the negative complexity-stability
relationships, but it the presence of response diversity and resulting
asynchrony in species fluctuations which overrides the negative effects of
complexity on stability-richness relationship. Response diversity might then one
mechanism that explains why so much positive relationship between stability and
species richness are found in empirical settings while food-web theoretical
model that do not often include environmental stochasticity and response
diversity finds negative complexity stability relationships
[@ives_stability_2007].

# Conclusion and perspectives

Our study presents the effects of food-web structure on temporal community
stability and its partition into population stability and asynchrony in complex
food-web, added to two pressing stressors in the context of global change, the
strength of environmental variability and the level of response diversity of
species facing this environmental variability. Our results confirm the results
of previous theoretical models modelling small food-web modules or food-chain,
as we report that average trophic level, omnivory, predator-prey size ratio, and
functional response of type III increase temporal community stability by
stabilising population fluctuations. In turn, our study confirm that species
richness increase temporal community stability by increasing asynchrony due to
portfolio effects and thanks to response diversity. Our study reports that
connectance and average interaction strength have little effects on overall
community stability while they have a strong effect on the sign of the community
stability-richness relationship, in accordance with the complexity-stability
paradox, but only in the absence of response diversity. Overall, it suggests that
high response diversity of species facing environmental stochasticity, on top of
being the essential ingredient driving positive stability-richness relationship,
also cancel the negative effects of food-web complexity on stability-richness
relationship.

Our results highlight that population stability is the primary driver of
community stability. In the context of global changes in which environmental
variability is expected to increase tremendously, we expect to see a strong
decrease in community stability, despite high response diversity. But ongoing
biotic homogenization [@olden_homogocene_2018; @muthukrishnan_invasive_2020;
@wang_biotic_2021] is also suggested to decrease response diversity
[@wang_biotic_2021; @su_human_2021], which in turn is expected to further
decrease stability despite high species richness, even more in complex food-webs
having high interaction strength, high connectance and high average trophic
level. Then, our study reinforces that the restoration of ecosystems along with
a restoration of landscape complexity that provides refuges from extreme
environment are key to maintain temporal stability of ecological communities
[@ryser_landscape_2021].
